---
title: "Results"
output: 
  html_document: 
    toc: true
    toc_float: true
    keep_md: true
    number_sections: true
date: "2025-04-24"
---
Return to Model README: [README](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/README.md)

# Exploring Metapopulation Capacity
## The independent effects of habitat cover on metapopulation capacity

Experimental Design: [The independent effects of habitat cover on metapopulation capacity - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-independent-effects-of-habitat-cover-on-metapopulation-capacity)

Changing the size of habitat patches, without changing the distance between the habitat patches gives the following relationship between metapopulation capacity and habitat cover:
```{r area_landscapes_results, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300}


rm(list = ls())

library(tidyverse)
library(patchwork)


# Generate and Plot Habitat Patches ---------------------------------------


# Define 10 equally spaced habitat patches
points_10 <- data.frame(
  x = seq(0, 100, length.out = 10),
  y = seq(0, 100, length.out = 10)
)

# Define 5 equally spaced habitat patches
points_5 <- data.frame(
  x = seq(0, 100, length.out = 5),
  y = seq(0, 100, length.out = 5)
)

# Define 2 equally spaced habitat patches
points_2 <- data.frame(
  x = seq(0, 100, length.out = 2),
  y = seq(0, 100, length.out = 2)
)



# Calculate Metapopulation Capacity ---------------------------------------

# Set alpha
alpha <- 0.05  # change as needed

distance_matrix_2 <- as.matrix(dist(points_2))
distance_matrix_5 <- as.matrix(dist(points_5))
distance_matrix_10 <- as.matrix(dist(points_10))

# Number of patches
n_patches_2 <- nrow(points_2)
n_patches_5 <- nrow(points_5)
n_patches_10 <- nrow(points_10)

# Create results dataframe
results_2 <- data.frame(
  habitat_area = numeric(),
  metapop_cap = numeric()
)
results_5 <- data.frame(
  habitat_area = numeric(),
  metapop_cap = numeric()
)
results_10 <- data.frame(
  habitat_area = numeric(),
  metapop_cap = numeric()
)


# Loop from habitat area 1 to 500 for 2 patches
for (a in 1:500) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_2, ncol = n_patches_2)
  
  dispersal_matrix <- exp(-alpha * distance_matrix_2)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # update results
  results_2 <- rbind(results_2, data.frame(habitat_area = a, metapop_cap = leading_eig))
}


# Loop from habitat area 1 to 500 for 5 patches
for (a in 1:500) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_5, ncol = n_patches_5)
  
  dispersal_matrix <- exp(-alpha * distance_matrix_5)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # update results
  results_5 <- rbind(results_5, data.frame(habitat_area = a, metapop_cap = leading_eig))
}



# Loop from habitat area 1 to 500 for 10 patches
for (a in 1:500) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_10, ncol = n_patches_10)
  
  
  dispersal_matrix <- exp(-alpha * distance_matrix_10)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix# Step 3: Connectivity matrix = area * dispersal
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # update results
  results_10 <- rbind(results_10, data.frame(habitat_area = a, metapop_cap = leading_eig))
}







# Plot Graphs -------------------------------------------------------------

results_2$habitat_cover <- ((results_2$habitat_area * 2)/(100*100)) *100
results_5$habitat_cover <- ((results_5$habitat_area * 5)/(100*100)) *100
results_10$habitat_cover <- ((results_10$habitat_area * 10)/(100*100)) *100


# Plot Metapopulation Capacity

mp_1 <- ggplot(results_2, aes(x = habitat_cover, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_2 <- ggplot(results_5, aes(x = habitat_cover, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_3 <- ggplot(results_10, aes(x = habitat_cover, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )
  

mp_1 / mp_2 / mp_3


```

The log of metapopulation capacity plotted against habitat cover gives the following results:

```{r area_landscapes_log_results, echo=FALSE, fig.height=8,  message=FALSE, warning=FALSE, dpi=300}


# Plot log metapopulation capacity

log_mp_1 <- ggplot(results_2, aes(x = habitat_cover, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_2 <- ggplot(results_5, aes(x = habitat_cover, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_3 <- ggplot(results_10, aes(x = habitat_cover, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )


log_mp_1 / log_mp_2 / log_mp_3



```

## The independent effects of inter-patch distance on metapopulation capacity 

Experimental Design: [The independent effects of habitat cover on metapopulation capacity - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-independent-effects-of-inter-patch-distance-on-metapopulation-capacity)

Scaling the dispersal matrix whilst keeping the habitat area constant gives the following relationship between metapopulation capacity and mean distance between habitat patches.
```{r dispersal_landscapes_results, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300}



rm(list = ls())

library(tidyverse)
library(patchwork)


# Generate and Plot Habitat Patches ---------------------------------------


# Define 10 equally spaced habitat patches
points_10 <- data.frame(
  x = seq(0, 100, length.out = 10),
  y = seq(0, 100, length.out = 10)
)

# Define 5 equally spaced habitat patches
points_5 <- data.frame(
  x = seq(0, 100, length.out = 5),
  y = seq(0, 100, length.out = 5)
)

# Define 2 equally spaced habitat patches
points_2 <- data.frame(
  x = seq(0, 100, length.out = 2),
  y = seq(0, 100, length.out = 2)
)



# Calculate Metapopulation Capacity ---------------------------------------

# Set alpha
alpha <- 0.05  

# Number of patches
n_patches_2 <- nrow(points_2)
n_patches_5 <- nrow(points_5)
n_patches_10 <- nrow(points_10)

distance_matrix_2 <- as.matrix(dist(points_2))
distance_matrix_5 <- as.matrix(dist(points_5))
distance_matrix_10 <- as.matrix(dist(points_10))

# Create results dataframe
results_2 <- data.frame(
  mean_distance = numeric(),
  metapop_cap = numeric()
)
results_5 <- data.frame(
  mean_distance = numeric(),
  metapop_cap = numeric()
)
results_10 <- data.frame(
  mean_distance = numeric(),
  metapop_cap = numeric()
)

# set area to 100 units
a <- 100

# Loop from distance multiplier of 1 to 500
for (d in seq(1, 5, by = 0.01)) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_2, ncol = n_patches_2)
  
  scaled_distance_matrix <- distance_matrix_2 *d
  
  dispersal_matrix <- exp(-alpha * scaled_distance_matrix)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # calculate mean distance
  dist_vector <- scaled_distance_matrix[lower.tri(scaled_distance_matrix)]
  mean_distance <- mean(dist_vector)
  
  # update results
  results_2 <- rbind(results_2, data.frame(mean_distance = mean_distance, metapop_cap = leading_eig))
}


# Loop from distance multiplier of 1 to 500
for (d in seq(1, 5, by = 0.01)) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_5, ncol = n_patches_5)
  
  distance_matrix <- distance_matrix_5 *d
  
  dispersal_matrix <- exp(-alpha * distance_matrix)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # calculate mean distance
  dist_vector <- distance_matrix[lower.tri(distance_matrix)]
  mean_distance <- mean(dist_vector)
  
  # update results
  results_5 <- rbind(results_5, data.frame(mean_distance = mean_distance, metapop_cap = leading_eig))
}




# Loop from distance multiplier of 1 to 500
for (d in seq(1, 5, by = 0.01)) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_10, ncol = n_patches_10)
  
  distance_matrix <- distance_matrix_10 *d
  
  dispersal_matrix <- exp(-alpha * distance_matrix)
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # calculate mean distance
  dist_vector <- distance_matrix[lower.tri(distance_matrix)]
  mean_distance <- mean(dist_vector)
  
  # update results
  results_10 <- rbind(results_10, data.frame(mean_distance = mean_distance, metapop_cap = leading_eig))
}







# Plot Graphs -------------------------------------------------------------


# Plot Metapopulation Capacity

mp_1 <- ggplot(results_2, aes(x = mean_distance, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_2 <- ggplot(results_5, aes(x = mean_distance, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_3 <- ggplot(results_10, aes(x = mean_distance, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )


mp_1 / mp_2 / mp_3

```


The log of metapopulation capacity plotted against mean distance between patches gives the following results:

```{r distance_landscapes_log_results, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300}

# Plot log metapopulation capacity

log_mp_1 <- ggplot(results_2, aes(x = mean_distance, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_2 <- ggplot(results_5, aes(x = mean_distance, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_3 <- ggplot(results_10, aes(x = mean_distance, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )


log_mp_1 / log_mp_2 / log_mp_3



```

## The effects of habitat configuration on patch distances and areas

Experimental Design: [The effects of habitat configuration on patch distances and areas - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-effects-of-habitat-configuration-on-patch-distances-and-areas)

```{r config_distance_results, echo=FALSE, message=FALSE, warning=FALSE, dpi=300}

rm(list = ls())

library(tidyverse)
library(cowplot)
library(patchwork)

df <- read.csv("Results_Folder/Exploring Metapopulation Capacity/Configuration Distance Relationship/config_dist_summary.csv")

p1 <- ggplot(df, aes(x = as.factor(p), y = edge_density)) +
  geom_boxplot() +
  labs(x = "p", y = "Edge Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p2 <- ggplot(df, aes(x = as.factor(p), y = mean_distance )) +
  geom_boxplot() +
  labs(x = "p", y = "Mean inter-patch Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p3 <- ggplot(df, aes(x = as.factor(p), y = mean_nearest_distance )) +
  geom_boxplot() +
  labs(x = "p", y = "Mean nearest-neighbour Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p4 <- ggplot(df, aes(x = as.factor(p), y = num_patches )) +
  geom_boxplot() +
  labs(x = "p", y = "Number of Patches") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p5 <- ggplot(df, aes(x = as.factor(p), y = mean_patch_area )) +
  geom_boxplot() +
  labs(x = "p", y = "Mean Habitat Patch Area") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p1 + p2 + p3 + p4 + p5


```


## The effects of habitat cover on patch distances and areas

Experimental Design: [The effects of habitat cover on patch distances and areas - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-effects-of-habitat-cover-on-patch-distances-and-areas)

```{r cover_distance_results, echo=FALSE, message=FALSE, warning=FALSE, dpi=300}


rm(list = ls())

library(tidyverse)
library(cowplot)
library(patchwork)

df <- read.csv("Results_Folder/Exploring Metapopulation Capacity/Cover Distance Relationship/cover_dist_summary.csv")

p1 <- ggplot(df, aes(x = as.factor(ai), y = edge_density)) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Edge Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p2 <- ggplot(df, aes(x = as.factor(ai), y = mean_distance )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Mean inter-patch Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p3 <- ggplot(df, aes(x = as.factor(ai), y = mean_nearest_distance )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Mean nearest-neighbour Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p4 <- ggplot(df, aes(x = as.factor(ai), y = num_patches )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Number of Patches") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p5 <- ggplot(df, aes(x = as.factor(ai), y = mean_patch_area )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Mean Habitat Patch Area") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p1 + p2 + p3 + p4 + p5


```


# The effect of Matrix Quality on Metapopulation Capacity and Persistence

## The effect of Matrix Quality on Metapopulation Capacity

```{r matrix_quality_data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list=ls())

library(tidyverse)

# Import Data -------------------------------------------------------------

# High aggregation group
ha_vhd <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_vhd.csv")
ha_hd  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_hd.csv")
ha_md  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_md.csv")
ha_ld  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_ld.csv")
ha_vld <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_vld.csv")

# Medium aggregation group
ma_vhd <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_vhd.csv")
ma_hd  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_hd.csv")
ma_md  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_md.csv")
ma_ld  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_ld.csv")
ma_vld <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_vld.csv")

# Low aggregation group
la_vhd <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_vhd.csv")
la_hd  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_hd.csv")
la_md  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_md.csv")
la_ld  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_ld.csv")
la_vld <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_vld.csv")






# Plot Metapopulation Capacities ------------------------------------------

library(ggplot2)
library(grid)
library(patchwork)

# Function for plotting graphs

plot_metapop_vs_cover <- function(df, df_name) {
  p <- ggplot(df, aes(x = habitat_cover, y = log(metapop_cap), colour = movement)) +
    geom_point() +
    geom_smooth(method = "gam", se = FALSE) +
    labs(
      x = "Habitat Cover",
      y = "log Metapopulation Capacity",
      color = "Movement Factor"
    ) +
    theme_minimal()
  
  # Create name like ha_vhd_plot and assign plot object to it in global environment
  plot_name <- paste0(df_name, "_plot")
  assign(plot_name, p, envir = .GlobalEnv)
}


# Function for dispersal labels

dispersal_label <- function(df_name) {
  level_code <- strsplit(df_name, "_")[[1]][2]
  switch(
    level_code,
    "vhd" = "Very High Dispersal",
    "hd" = "High Dispersal",
    "md" = "Medium Dispersal",
    "ld" = "Low Dispersal",
    "vld" = "Very Low Dispersal",
    level_code  # fallback
  )
}

```


### Landscapes with high levels of habitat aggregation

High levels of habitat aggregation with a maximum dispersal distance increase of 50%
```{r matrix_quality_ha_50, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}
# Plot for 50% increase ---------------------------------------------------

# Filter for 50% increases
ha_vhd_50 <- ha_vhd[grepl("50$", as.character(ha_hd$movement)), ]
ha_hd_50 <- ha_hd[grepl("50$", as.character(ha_hd$movement)), ]
ha_md_50 <- ha_md[grepl("50$", as.character(ha_hd$movement)), ]
ha_ld_50 <- ha_ld[grepl("50$", as.character(ha_hd$movement)), ]
ha_vld_50 <- ha_vld[grepl("50$", as.character(ha_hd$movement)), ]


plot_metapop_vs_cover(ha_vhd_50, "ha_vhd_50")
plot_metapop_vs_cover(ha_hd_50, "ha_hd_50")
plot_metapop_vs_cover(ha_md_50, "ha_md_50")
plot_metapop_vs_cover(ha_ld_50, "ha_ld_50")
plot_metapop_vs_cover(ha_vld_50, "ha_vld_50")



# Create labeled plots
ha_vhd_labeled_50 <- ha_vhd_50_plot + ggtitle(dispersal_label("ha_vhd"))
ha_hd_labeled_50 <- ha_hd_50_plot + ggtitle(dispersal_label("ha_hd"))
ha_md_labeled_50 <- ha_md_50_plot + ggtitle(dispersal_label("ha_md"))
ha_ld_labeled_50 <- ha_ld_50_plot + ggtitle(dispersal_label("ha_ld"))
ha_vld_labeled_50 <- ha_vld_50_plot + ggtitle(dispersal_label("ha_vld"))


# Combine them with shared legend
combined_plot <- (
  ha_vhd_labeled_50 + ha_hd_labeled_50 + ha_md_labeled_50
) / (
  ha_ld_labeled_50 + ha_vld_labeled_50 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot

```

High levels of habitat aggregation with a maximum dispersal distance increase of 500%
```{r matrix_quality_ha_500, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 500% increase ---------------------------------------------------

# Filter for 500% increases
ha_vhd_500 <- ha_vhd[grepl("500$", as.character(ha_hd$movement)), ]
ha_hd_500 <- ha_hd[grepl("500$", as.character(ha_hd$movement)), ]
ha_md_500 <- ha_md[grepl("500$", as.character(ha_hd$movement)), ]
ha_ld_500 <- ha_ld[grepl("500$", as.character(ha_hd$movement)), ]
ha_vld_500 <- ha_vld[grepl("500$", as.character(ha_hd$movement)), ]


plot_metapop_vs_cover(ha_vhd_500, "ha_vhd_500")
plot_metapop_vs_cover(ha_hd_500, "ha_hd_500")
plot_metapop_vs_cover(ha_md_500, "ha_md_500")
plot_metapop_vs_cover(ha_ld_500, "ha_ld_500")
plot_metapop_vs_cover(ha_vld_500, "ha_vld_500")



# Create labeled plots
ha_vhd_labeled_500 <- ha_vhd_500_plot + ggtitle(dispersal_label("ha_vhd"))
ha_hd_labeled_500 <- ha_hd_500_plot + ggtitle(dispersal_label("ha_hd"))
ha_md_labeled_500 <- ha_md_500_plot + ggtitle(dispersal_label("ha_md"))
ha_ld_labeled_500 <- ha_ld_500_plot + ggtitle(dispersal_label("ha_ld"))
ha_vld_labeled_500 <- ha_vld_500_plot + ggtitle(dispersal_label("ha_vld"))


# Combine them with shared legend
combined_plot <- (
  ha_vhd_labeled_500 + ha_hd_labeled_500 + ha_md_labeled_500
) / (
  ha_ld_labeled_500 + ha_vld_labeled_500 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot

```

High levels of habitat aggregation with a maximum dispersal distance increase of 5000%
```{r matrix_quality_ha_5000, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 5000% increase ---------------------------------------------------

# Filter for 5000% increases
ha_vhd_5000 <- ha_vhd[grepl("5000$", as.character(ha_hd$movement)), ]
ha_hd_5000 <- ha_hd[grepl("5000$", as.character(ha_hd$movement)), ]
ha_md_5000 <- ha_md[grepl("5000$", as.character(ha_hd$movement)), ]
ha_ld_5000 <- ha_ld[grepl("5000$", as.character(ha_hd$movement)), ]
ha_vld_5000 <- ha_vld[grepl("5000$", as.character(ha_hd$movement)), ]


plot_metapop_vs_cover(ha_vhd_5000, "ha_vhd_5000")
plot_metapop_vs_cover(ha_hd_5000, "ha_hd_5000")
plot_metapop_vs_cover(ha_md_5000, "ha_md_5000")
plot_metapop_vs_cover(ha_ld_5000, "ha_ld_5000")
plot_metapop_vs_cover(ha_vld_5000, "ha_vld_5000")



# Create labeled plots
ha_vhd_labeled_5000 <- ha_vhd_5000_plot + ggtitle(dispersal_label("ha_vhd"))
ha_hd_labeled_5000 <- ha_hd_5000_plot + ggtitle(dispersal_label("ha_hd"))
ha_md_labeled_5000 <- ha_md_5000_plot + ggtitle(dispersal_label("ha_md"))
ha_ld_labeled_5000 <- ha_ld_5000_plot + ggtitle(dispersal_label("ha_ld"))
ha_vld_labeled_5000 <- ha_vld_5000_plot + ggtitle(dispersal_label("ha_vld"))


# Combine them with shared legend
combined_plot <- (
  ha_vhd_labeled_5000 + ha_hd_labeled_5000 + ha_md_labeled_5000
) / (
  ha_ld_labeled_5000 + ha_vld_labeled_5000 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot


```

### Landscapes with medium levels of habitat aggregation

Medium levels of habitat aggregation with a maximum dispersal distance increase of 50%
```{r matrix_quality_ma_50, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 50% increase ---------------------------------------------------

# Filter for 50% increases
ma_vhd_50 <- ma_vhd[grepl("50$", as.character(ma_hd$movement)), ]
ma_hd_50 <- ma_hd[grepl("50$", as.character(ma_hd$movement)), ]
ma_md_50 <- ma_md[grepl("50$", as.character(ma_hd$movement)), ]
ma_ld_50 <- ma_ld[grepl("50$", as.character(ma_hd$movement)), ]
ma_vld_50 <- ma_vld[grepl("50$", as.character(ma_hd$movement)), ]


plot_metapop_vs_cover(ma_vhd_50, "ma_vhd_50")
plot_metapop_vs_cover(ma_hd_50, "ma_hd_50")
plot_metapop_vs_cover(ma_md_50, "ma_md_50")
plot_metapop_vs_cover(ma_ld_50, "ma_ld_50")
plot_metapop_vs_cover(ma_vld_50, "ma_vld_50")



# Create labeled plots
ma_vhd_labeled_50 <- ma_vhd_50_plot + ggtitle(dispersal_label("ma_vhd"))
ma_hd_labeled_50 <- ma_hd_50_plot + ggtitle(dispersal_label("ma_hd"))
ma_md_labeled_50 <- ma_md_50_plot + ggtitle(dispersal_label("ma_md"))
ma_ld_labeled_50 <- ma_ld_50_plot + ggtitle(dispersal_label("ma_ld"))
ma_vld_labeled_50 <- ma_vld_50_plot + ggtitle(dispersal_label("ma_vld"))


# Combine them with smared legend
combined_plot <- (
  ma_vhd_labeled_50 + ma_hd_labeled_50 + ma_md_labeled_50
) / (
  ma_ld_labeled_50 + ma_vld_labeled_50 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot


```

Medium levels of habitat aggregation with a maximum dispersal distance increase of 500%
```{r matrix_quality_ma_500, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}


# Plot for 500% increase ---------------------------------------------------

# Filter for 500% increases
ma_vhd_500 <- ma_vhd[grepl("500$", as.character(ma_hd$movement)), ]
ma_hd_500 <- ma_hd[grepl("500$", as.character(ma_hd$movement)), ]
ma_md_500 <- ma_md[grepl("500$", as.character(ma_hd$movement)), ]
ma_ld_500 <- ma_ld[grepl("500$", as.character(ma_hd$movement)), ]
ma_vld_500 <- ma_vld[grepl("500$", as.character(ma_hd$movement)), ]


plot_metapop_vs_cover(ma_vhd_500, "ma_vhd_500")
plot_metapop_vs_cover(ma_hd_500, "ma_hd_500")
plot_metapop_vs_cover(ma_md_500, "ma_md_500")
plot_metapop_vs_cover(ma_ld_500, "ma_ld_500")
plot_metapop_vs_cover(ma_vld_500, "ma_vld_500")



# Create labeled plots
ma_vhd_labeled_500 <- ma_vhd_500_plot + ggtitle(dispersal_label("ma_vhd"))
ma_hd_labeled_500 <- ma_hd_500_plot + ggtitle(dispersal_label("ma_hd"))
ma_md_labeled_500 <- ma_md_500_plot + ggtitle(dispersal_label("ma_md"))
ma_ld_labeled_500 <- ma_ld_500_plot + ggtitle(dispersal_label("ma_ld"))
ma_vld_labeled_500 <- ma_vld_500_plot + ggtitle(dispersal_label("ma_vld"))


# Combine them with smared legend
combined_plot <- (
  ma_vhd_labeled_500 + ma_hd_labeled_500 + ma_md_labeled_500
) / (
  ma_ld_labeled_500 + ma_vld_labeled_500 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot



```

Medium levels of habitat aggregation with a maximum dispersal distance increase of 5000%
```{r matrix_quality_ma_5000, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}


# Plot for 5000% increase ---------------------------------------------------

# Filter for 5000% increases
ma_vhd_5000 <- ma_vhd[grepl("5000$", as.character(ma_hd$movement)), ]
ma_hd_5000 <- ma_hd[grepl("5000$", as.character(ma_hd$movement)), ]
ma_md_5000 <- ma_md[grepl("5000$", as.character(ma_hd$movement)), ]
ma_ld_5000 <- ma_ld[grepl("5000$", as.character(ma_hd$movement)), ]
ma_vld_5000 <- ma_vld[grepl("5000$", as.character(ma_hd$movement)), ]


plot_metapop_vs_cover(ma_vhd_5000, "ma_vhd_5000")
plot_metapop_vs_cover(ma_hd_5000, "ma_hd_5000")
plot_metapop_vs_cover(ma_md_5000, "ma_md_5000")
plot_metapop_vs_cover(ma_ld_5000, "ma_ld_5000")
plot_metapop_vs_cover(ma_vld_5000, "ma_vld_5000")



# Create labeled plots
ma_vhd_labeled_5000 <- ma_vhd_5000_plot + ggtitle(dispersal_label("ma_vhd"))
ma_hd_labeled_5000 <- ma_hd_5000_plot + ggtitle(dispersal_label("ma_hd"))
ma_md_labeled_5000 <- ma_md_5000_plot + ggtitle(dispersal_label("ma_md"))
ma_ld_labeled_5000 <- ma_ld_5000_plot + ggtitle(dispersal_label("ma_ld"))
ma_vld_labeled_5000 <- ma_vld_5000_plot + ggtitle(dispersal_label("ma_vld"))


# Combine them with smared legend
combined_plot <- (
  ma_vhd_labeled_5000 + ma_hd_labeled_5000 + ma_md_labeled_5000
) / (
  ma_ld_labeled_5000 + ma_vld_labeled_5000 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot


```



### Landscapes with medium levels of habitat aggregation

Low levels of habitat aggregation with a maximum dispersal distance increase of 50%
```{r matrix_quality_la_50, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Low Aggregation Plots --------------------------------------------------


# Plot for 50% increase ---------------------------------------------------

# Filter for 50% increases
la_vhd_50 <- la_vhd[grepl("50$", as.character(la_hd$movement)), ]
la_hd_50 <- la_hd[grepl("50$", as.character(la_hd$movement)), ]
la_md_50 <- la_md[grepl("50$", as.character(la_hd$movement)), ]
la_ld_50 <- la_ld[grepl("50$", as.character(la_hd$movement)), ]
la_vld_50 <- la_vld[grepl("50$", as.character(la_hd$movement)), ]


plot_metapop_vs_cover(la_vhd_50, "la_vhd_50")
plot_metapop_vs_cover(la_hd_50, "la_hd_50")
plot_metapop_vs_cover(la_md_50, "la_md_50")
plot_metapop_vs_cover(la_ld_50, "la_ld_50")
plot_metapop_vs_cover(la_vld_50, "la_vld_50")



# Create labeled plots
la_vhd_labeled_50 <- la_vhd_50_plot + ggtitle(dispersal_label("la_vhd"))
la_hd_labeled_50 <- la_hd_50_plot + ggtitle(dispersal_label("la_hd"))
la_md_labeled_50 <- la_md_50_plot + ggtitle(dispersal_label("la_md"))
la_ld_labeled_50 <- la_ld_50_plot + ggtitle(dispersal_label("la_ld"))
la_vld_labeled_50 <- la_vld_50_plot + ggtitle(dispersal_label("la_vld"))


# Combine them with slared legend
combined_plot <- (
  la_vhd_labeled_50 + la_hd_labeled_50 + la_md_labeled_50
) / (
  la_ld_labeled_50 + la_vld_labeled_50 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot



```

Low levels of habitat aggregation with a maximum dispersal distance increase of 500%
```{r matrix_quality_la_500, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 500% increase ---------------------------------------------------

# Filter for 500% increases
la_vhd_500 <- la_vhd[grepl("500$", as.character(la_hd$movement)), ]
la_hd_500 <- la_hd[grepl("500$", as.character(la_hd$movement)), ]
la_md_500 <- la_md[grepl("500$", as.character(la_hd$movement)), ]
la_ld_500 <- la_ld[grepl("500$", as.character(la_hd$movement)), ]
la_vld_500 <- la_vld[grepl("500$", as.character(la_hd$movement)), ]


plot_metapop_vs_cover(la_vhd_500, "la_vhd_500")
plot_metapop_vs_cover(la_hd_500, "la_hd_500")
plot_metapop_vs_cover(la_md_500, "la_md_500")
plot_metapop_vs_cover(la_ld_500, "la_ld_500")
plot_metapop_vs_cover(la_vld_500, "la_vld_500")



# Create labeled plots
la_vhd_labeled_500 <- la_vhd_500_plot + ggtitle(dispersal_label("la_vhd"))
la_hd_labeled_500 <- la_hd_500_plot + ggtitle(dispersal_label("la_hd"))
la_md_labeled_500 <- la_md_500_plot + ggtitle(dispersal_label("la_md"))
la_ld_labeled_500 <- la_ld_500_plot + ggtitle(dispersal_label("la_ld"))
la_vld_labeled_500 <- la_vld_500_plot + ggtitle(dispersal_label("la_vld"))


# Combine them with slared legend
combined_plot <- (
  la_vhd_labeled_500 + la_hd_labeled_500 + la_md_labeled_500
) / (
  la_ld_labeled_500 + la_vld_labeled_500 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot




```

Low levels of habitat aggregation with a maximum dispersal distance increase of 5000%
```{r matrix_quality_la_5000, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}


# Plot for 5000% increase ---------------------------------------------------

# Filter for 5000% increases
la_vhd_5000 <- la_vhd[grepl("5000$", as.character(la_hd$movement)), ]
la_hd_5000 <- la_hd[grepl("5000$", as.character(la_hd$movement)), ]
la_md_5000 <- la_md[grepl("5000$", as.character(la_hd$movement)), ]
la_ld_5000 <- la_ld[grepl("5000$", as.character(la_hd$movement)), ]
la_vld_5000 <- la_vld[grepl("5000$", as.character(la_hd$movement)), ]


plot_metapop_vs_cover(la_vhd_5000, "la_vhd_5000")
plot_metapop_vs_cover(la_hd_5000, "la_hd_5000")
plot_metapop_vs_cover(la_md_5000, "la_md_5000")
plot_metapop_vs_cover(la_ld_5000, "la_ld_5000")
plot_metapop_vs_cover(la_vld_5000, "la_vld_5000")



# Create labeled plots
la_vhd_labeled_5000 <- la_vhd_5000_plot + ggtitle(dispersal_label("la_vhd"))
la_hd_labeled_5000 <- la_hd_5000_plot + ggtitle(dispersal_label("la_hd"))
la_md_labeled_5000 <- la_md_5000_plot + ggtitle(dispersal_label("la_md"))
la_ld_labeled_5000 <- la_ld_5000_plot + ggtitle(dispersal_label("la_ld"))
la_vld_labeled_5000 <- la_vld_5000_plot + ggtitle(dispersal_label("la_vld"))


# Combine them with slared legend
combined_plot <- (
  la_vhd_labeled_5000 + la_hd_labeled_5000 + la_md_labeled_5000
) / (
  la_ld_labeled_5000 + la_vld_labeled_5000 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot

```


## The effect of matrix quality on metapopulation persistence (Medium Colonisation/ Medium Extinction)

```{r gam_differences, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, dpi=300, fig.height=8}
rm(list = ls())

library(tidyverse)

diff_results <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/GAM/gam_differences.csv")

# Plot GAM differences ----------------------------------------------------

plot_diff_vs_habitat <- function(diff_df, source_group, type, mag, delta_prefix = "delta_") {
  df_sub <- diff_df[diff_df$source_1 %in% source_group & grepl(delta_prefix, diff_df$source_2), ]
  
  # Human-readable labels
  type_label <- switch(type,
                       "ha" = "Low Aggregation",
                       "ma" = "Medium Aggregation",
                       "la" = "High Aggregation",
                       type)
  mag_label <- paste0(mag, "% Movement Increase")
  
  ggplot(df_sub, aes(x = habitat_cover, y = abs_diff, color = source_2)) +
    geom_line(alpha = 0.8) +
    labs(
      title = paste("Absolute Difference in Metapopulation Capacity –", type_label, "–", mag_label),
      x = "Habitat Cover",
      y = "Absolute Difference",
      color = "Delta Scenario"
    ) +
    theme_minimal() +
    facet_wrap(~ source_1, ncol = 3, scales = "free_y") +
    theme(legend.position = "bottom")
}


# Extract unique source_1 names
unique_sources <- unique(diff_results$source_1)

# Get combinations of aggregation type and movement ability
types <- c("ha", "ma", "la")
magnitudes <- c("50", "500", "5000")

plot_list <- list()

for (type in types) {
  for (mag in magnitudes) {
    pattern <- paste0("^", type, "_.*_", mag, "_")
    group_sources <- unique_sources[grepl(pattern, unique_sources)]
    
    if (length(group_sources) == 0) next
    
    p <- plot_diff_vs_habitat(diff_results, source_group = group_sources, type = type, mag = mag)
    print(p)
    plot_list[[paste0(type, "_", mag)]] <- p
  }
}
```







## The effect of matrix quality on metapopulation persistence (High Colonisation/ Low Extinction)

```{r gam_differences_h_c_l_e, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300, paged.print=FALSE}

rm(list = ls())

library(tidyverse)
library(mgcv)

# Generate habitat covers for predicting data
habitat_seq <- data.frame(habitat_cover = seq(0, 45, by = 0.1))

# Prepare Model Results Data ----------------------------------------------

all_results <- readRDS("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/GAM/model_gam_results_clean")


# Prepare Delta Data ------------------------------------------------------


# Define movement responses
# movement can be 'linear', 'concave', or 'convex'
movements <- c('linear', 'concave', 'convex')

# Define maximum movement increases (%)
movement_abilities <- c(50, 500, 5000)

# Create grid of movement combinations
movement_combos <- expand.grid(movement = movements, movement_ability = movement_abilities)

# Create a new row for null movement
new_row <- data.frame(movement = "none", movement_ability = 0)

# Add the new row to the grid
movement_combos <- rbind(movement_combos, new_row)
rm(new_row)

# Set movement as a character
movement_combos$movement <- as.character(movement_combos$movement)

# Create dataframe of slope constants for movement scenarios
m <- c("linear", "linear", "linear", "linear", "linear", "linear", "concave", "concave", "concave", "concave", "concave", "concave", "convex", "convex", "convex", "convex", "convex", "convex")
m_a <- c("10", "50", "100", "500", "1000", "5000", "10", "50", "100", "500", "1000", "5000", "10", "50", "100", "500", "1000", "5000")
s1 <- c(0.1, 0.5, 1, 5, 10, 50, 0.1, 0.5, 1, 5, 10, 50, 4.54e-6, 2.27e-5, 4.54e-5, 2.27e-4, 4.53e-4, 0.00227)
s2 <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 0.99999546, 0, 0, 0, 0, 0.99773)
a <- c(1, 1, 1, 1, 1, 1, 0.1, 0.5, 1, 5, 10, 50, 0, 1, 1, 1, 1, 0)

slopes <- data.frame(m, m_a, s1, s2, a)

# Create an empty list to store the results
c_multiplier_list <- list()

for (j in 1:nrow(movement_combos)) {
  movement <- movement_combos$movement[j]
  movement_ability <- movement_combos$movement_ability[j]
  
  # Matrix Effects ----------------------------------------------------------
  
  productivity <- 0.5
  
  landscape_cover <- seq(0, 45, by = 1)
  
  # Calculate agricultural yield
  yield <- productivity / (1 - (landscape_cover / 100))
  
  # Normalize yield to get relative yield
  relative_yield <- (yield - productivity) / (1 - productivity)
  
  
  
  # Functions for calculating alpha (dispersal distance)
  # Linear
  linear_permeability <- function(y, s, a) {
    disp_factor = (((s * (1 - y)) + a))
    return(disp_factor)
  }
  
  # Concave
  conc_permeability <- function(y, s, a) {
    disp_factor = (((-(y)^2 * s) + 1) + a)
    return(disp_factor)
  }
  
  # Convex
  conv_permeability <- function(y, s, s2, a) {
    disp_factor = (((exp(10 * (1 - y))) * s) + s2) + a
    return(disp_factor)
  }
  
  # Pick correct constants (slopes) for current dispersal scenario
  slope <- slopes %>% filter(m == movement & m_a == movement_ability)
  slope <- as.numeric(slope[3])
  slope_2 <- slopes %>% filter(m == movement & m_a == movement_ability)
  slope_2 <- as.numeric(slope_2[4])
  a <- slopes %>% filter(m == movement & m_a == movement_ability)
  a <- as.numeric(a[5])
  
  # Calculate dispersal multiplier
  if (movement == "linear") {
    dispersal_factor <- linear_permeability(y = relative_yield, s = slope, a = a)
  } else if (movement == "concave") {
    dispersal_factor <- conc_permeability(y = relative_yield, s = slope, a = a)
  } else if (movement == "convex") {
    dispersal_factor <- conv_permeability(y = relative_yield, s = slope, s2 = slope_2, a = a)
  } else if (movement == "none") {
    dispersal_factor <- 1
  } else {
    print("No matching movement")
  }
  
  c_intense <- 1
  
  # Create the dynamic variable name
  var_name <- paste("c_multiplier_", movement, "_", movement_ability, sep = "")
  
  # Calculate the value
  c_multiplier <- c_intense * dispersal_factor
  
  # Store the result in the list with dynamic variable names
  c_multiplier_list[[var_name]] <- c_multiplier  # Store it in the list
  
}



# Create a dataframe to hold the output
output_df <- data.frame(movement = character(0), habitat_cover = numeric(0), c_multiplier = numeric(0))

# Iterate through the list and combine with habitat_cover
for (var_name in names(c_multiplier_list)) {
  
  # Extract movement and movement_ability from the variable name
  split_name <- unlist(strsplit(var_name, "_"))
  movement <- split_name[3]
  movement_ability <- as.numeric(split_name[4])
  
  # Habitat cover values
  habitat_cover <- seq(0, 45, by = 1)  
  
  # Get the value of the c_multiplier for this specific variable
  c_multiplier_value <- c_multiplier_list[[var_name]]
  
  # Create a temporary dataframe for this combination
  temp_df <- data.frame(movement = rep(movement, length(habitat_cover)),
                        movement_ability = rep(movement_ability, length(habitat_cover)),
                        habitat_cover = habitat_cover,
                        c_multiplier = rep(c_multiplier_value, length(habitat_cover)))
  
  # Add to the main dataframe
  output_df <- rbind(output_df, temp_df)
}



delta_df <- output_df

delta_df$c_intense <- 0.5

delta_df$c <- delta_df$c_intense * delta_df$c_multiplier

delta_df$e <- 0.01

delta_df$delta <- delta_df$e/delta_df$c






process_delta_dataset <- function(df, dataset_name, habitat_grid) {
  # Get unique combinations of movement and movement_ability
  combos <- unique(df[, c("movement", "movement_ability")])
  
  results <- list()
  
  for (i in 1:nrow(combos)) {
    mov <- combos$movement[i]
    abil <- combos$movement_ability[i]
    
    # Filter data for this combination
    df_subset <- subset(df,
                        movement == mov &
                          movement_ability == abil &
                          !is.na(delta) &
                          is.finite(delta) &
                          !is.na(habitat_cover))
    
    if (nrow(df_subset) < 10) {
      warning(paste(dataset_name, mov, abil, "has too few rows, skipping GAM fit"))
      next
    }
    
    # Fit GAM: delta ~ s(habitat_cover)
    gam_model <- gam(delta ~ s(habitat_cover), data = df_subset)
    
    # Predict on grid
    pred_vals <- predict(gam_model, newdata = habitat_grid)
    pred_df <- data.frame(habitat_cover = habitat_grid$habitat_cover, pred = pred_vals)
    
    # Generate object names
    safe_mov <- gsub("[^a-zA-Z0-9]", "", tolower(mov))
    pred_name <- paste0(dataset_name, "_", safe_mov, "_", abil, "_pred")
    gam_name  <- paste0(dataset_name, "_", safe_mov, "_", abil, "_gam")
    data_name <- paste0(dataset_name, "_", safe_mov, "_", abil)
    
    # Store
    results[[pred_name]] <- pred_df
    results[[gam_name]]  <- gam_model
    results[[data_name]] <- df_subset
  }
  
  return(results)
}

delta_results <- process_delta_dataset(df = delta_df, dataset_name = "delta", habitat_grid = habitat_seq)


# Calculate GAM differences -----------------------------------------------

compare_all_predicted_gams <- function(all_results, delta_preds) {
  all_comparisons <- list()
  
  # Filter only *_pred entries
  preds_main <- all_results[grepl("_pred$", names(all_results))]
  preds_delta <- delta_preds[grepl("_pred$", names(delta_preds))]
  
  for (name_main in names(preds_main)) {
    df1 <- preds_main[[name_main]]
    source1_clean <- gsub("_pred$", "", name_main)
    
    for (name_delta in names(preds_delta)) {
      df2 <- preds_delta[[name_delta]]
      source2_clean <- gsub("_pred$", "", name_delta)
      
      # Ensure matching habitat_cover grid
      if (!all.equal(df1$habitat_cover, df2$habitat_cover)) {
        warning(paste("Skipping mismatched habitat_cover between", name_main, "and", name_delta))
        next
      }
      
      # Compute difference and ratio on the original scale
      comparison_df <- data.frame(
        habitat_cover = df1$habitat_cover,
        source_1 = source1_clean,
        source_2 = source2_clean,
        abs_diff = df1$pred - df2$pred,
        abs_ratio = exp(df1$pred) / exp(df2$pred),
        stringsAsFactors = FALSE
      )
      
      all_comparisons[[paste0(source1_clean, "_vs_", source2_clean)]] <- comparison_df
    }
  }
  
  final_df <- do.call(rbind, all_comparisons)
  rownames(final_df) <- NULL
  return(final_df)
}


diff_results <- compare_all_predicted_gams(all_results, delta_results)


# Plot GAM differences ----------------------------------------------------

plot_diff_vs_habitat <- function(diff_df, source_group, type, mag, delta_prefix = "delta_") {
  df_sub <- diff_df[diff_df$source_1 %in% source_group & grepl(delta_prefix, diff_df$source_2), ]
  
  # Human-readable labels
  type_label <- switch(type,
                       "ha" = "Low Aggregation",
                       "ma" = "Medium Aggregation",
                       "la" = "High Aggregation",
                       type)
  mag_label <- paste0(mag, "% Movement Increase")
  
  ggplot(df_sub, aes(x = habitat_cover, y = abs_diff, color = source_2)) +
    geom_line(alpha = 0.8) +
    labs(
      title = paste("Absolute Difference in Metapopulation Capacity –", type_label, "–", mag_label),
      x = "Habitat Cover",
      y = "Absolute Difference",
      color = "Delta Scenario"
    ) +
    theme_minimal() +
    facet_wrap(~ source_1, ncol = 3, scales = "free_y") +
    theme(legend.position = "bottom")
}


# Extract unique source_1 names
unique_sources <- unique(diff_results$source_1)

# Get combinations of aggregation type and movement ability
types <- c("ha", "ma", "la")
magnitudes <- c("50", "500", "5000")

plot_list <- list()

for (type in types) {
  for (mag in magnitudes) {
    pattern <- paste0("^", type, "_.*_", mag, "_")
    group_sources <- unique_sources[grepl(pattern, unique_sources)]
    
    if (length(group_sources) == 0) next
    
    p <- plot_diff_vs_habitat(diff_results, source_group = group_sources, type = type, mag = mag)
    print(p)
    plot_list[[paste0(type, "_", mag)]] <- p
  }
}



```







## The effect of matrix quality on metapopulation persistence (Low Colonisation/ High Extinction)

```{r gam_differences_l_c_h_e, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300, paged.print=FALSE}

rm(list = ls())

library(tidyverse)
library(mgcv)

# Generate habitat covers for predicting data
habitat_seq <- data.frame(habitat_cover = seq(0, 45, by = 0.1))

# Prepare Model Results Data ----------------------------------------------

all_results <- readRDS("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/GAM/model_gam_results_clean")


# Prepare Delta Data ------------------------------------------------------


# Define movement responses
# movement can be 'linear', 'concave', or 'convex'
movements <- c('linear', 'concave', 'convex')

# Define maximum movement increases (%)
movement_abilities <- c(50, 500, 5000)

# Create grid of movement combinations
movement_combos <- expand.grid(movement = movements, movement_ability = movement_abilities)

# Create a new row for null movement
new_row <- data.frame(movement = "none", movement_ability = 0)

# Add the new row to the grid
movement_combos <- rbind(movement_combos, new_row)
rm(new_row)

# Set movement as a character
movement_combos$movement <- as.character(movement_combos$movement)

# Create dataframe of slope constants for movement scenarios
m <- c("linear", "linear", "linear", "linear", "linear", "linear", "concave", "concave", "concave", "concave", "concave", "concave", "convex", "convex", "convex", "convex", "convex", "convex")
m_a <- c("10", "50", "100", "500", "1000", "5000", "10", "50", "100", "500", "1000", "5000", "10", "50", "100", "500", "1000", "5000")
s1 <- c(0.1, 0.5, 1, 5, 10, 50, 0.1, 0.5, 1, 5, 10, 50, 4.54e-6, 2.27e-5, 4.54e-5, 2.27e-4, 4.53e-4, 0.00227)
s2 <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 0.99999546, 0, 0, 0, 0, 0.99773)
a <- c(1, 1, 1, 1, 1, 1, 0.1, 0.5, 1, 5, 10, 50, 0, 1, 1, 1, 1, 0)

slopes <- data.frame(m, m_a, s1, s2, a)

# Create an empty list to store the results
c_multiplier_list <- list()

for (j in 1:nrow(movement_combos)) {
  movement <- movement_combos$movement[j]
  movement_ability <- movement_combos$movement_ability[j]
  
  # Matrix Effects ----------------------------------------------------------
  
  productivity <- 0.5
  
  landscape_cover <- seq(0, 45, by = 1)
  
  # Calculate agricultural yield
  yield <- productivity / (1 - (landscape_cover / 100))
  
  # Normalize yield to get relative yield
  relative_yield <- (yield - productivity) / (1 - productivity)
  
  
  
  # Functions for calculating alpha (dispersal distance)
  # Linear
  linear_permeability <- function(y, s, a) {
    disp_factor = (((s * (1 - y)) + a))
    return(disp_factor)
  }
  
  # Concave
  conc_permeability <- function(y, s, a) {
    disp_factor = (((-(y)^2 * s) + 1) + a)
    return(disp_factor)
  }
  
  # Convex
  conv_permeability <- function(y, s, s2, a) {
    disp_factor = (((exp(10 * (1 - y))) * s) + s2) + a
    return(disp_factor)
  }
  
  # Pick correct constants (slopes) for current dispersal scenario
  slope <- slopes %>% filter(m == movement & m_a == movement_ability)
  slope <- as.numeric(slope[3])
  slope_2 <- slopes %>% filter(m == movement & m_a == movement_ability)
  slope_2 <- as.numeric(slope_2[4])
  a <- slopes %>% filter(m == movement & m_a == movement_ability)
  a <- as.numeric(a[5])
  
  # Calculate dispersal multiplier
  if (movement == "linear") {
    dispersal_factor <- linear_permeability(y = relative_yield, s = slope, a = a)
  } else if (movement == "concave") {
    dispersal_factor <- conc_permeability(y = relative_yield, s = slope, a = a)
  } else if (movement == "convex") {
    dispersal_factor <- conv_permeability(y = relative_yield, s = slope, s2 = slope_2, a = a)
  } else if (movement == "none") {
    dispersal_factor <- 1
  } else {
    print("No matching movement")
  }
  
  c_intense <- 1
  
  # Create the dynamic variable name
  var_name <- paste("c_multiplier_", movement, "_", movement_ability, sep = "")
  
  # Calculate the value
  c_multiplier <- c_intense * dispersal_factor
  
  # Store the result in the list with dynamic variable names
  c_multiplier_list[[var_name]] <- c_multiplier  # Store it in the list
  
}



# Create a dataframe to hold the output
output_df <- data.frame(movement = character(0), habitat_cover = numeric(0), c_multiplier = numeric(0))

# Iterate through the list and combine with habitat_cover
for (var_name in names(c_multiplier_list)) {
  
  # Extract movement and movement_ability from the variable name
  split_name <- unlist(strsplit(var_name, "_"))
  movement <- split_name[3]
  movement_ability <- as.numeric(split_name[4])
  
  # Habitat cover values
  habitat_cover <- seq(0, 45, by = 1)  
  
  # Get the value of the c_multiplier for this specific variable
  c_multiplier_value <- c_multiplier_list[[var_name]]
  
  # Create a temporary dataframe for this combination
  temp_df <- data.frame(movement = rep(movement, length(habitat_cover)),
                        movement_ability = rep(movement_ability, length(habitat_cover)),
                        habitat_cover = habitat_cover,
                        c_multiplier = rep(c_multiplier_value, length(habitat_cover)))
  
  # Add to the main dataframe
  output_df <- rbind(output_df, temp_df)
}



delta_df <- output_df

delta_df$c_intense <- 0.01

delta_df$c <- delta_df$c_intense * delta_df$c_multiplier

delta_df$e <- 0.5

delta_df$delta <- delta_df$e/delta_df$c






process_delta_dataset <- function(df, dataset_name, habitat_grid) {
  # Get unique combinations of movement and movement_ability
  combos <- unique(df[, c("movement", "movement_ability")])
  
  results <- list()
  
  for (i in 1:nrow(combos)) {
    mov <- combos$movement[i]
    abil <- combos$movement_ability[i]
    
    # Filter data for this combination
    df_subset <- subset(df,
                        movement == mov &
                          movement_ability == abil &
                          !is.na(delta) &
                          is.finite(delta) &
                          !is.na(habitat_cover))
    
    if (nrow(df_subset) < 10) {
      warning(paste(dataset_name, mov, abil, "has too few rows, skipping GAM fit"))
      next
    }
    
    # Fit GAM: delta ~ s(habitat_cover)
    gam_model <- gam(delta ~ s(habitat_cover), data = df_subset)
    
    # Predict on grid
    pred_vals <- predict(gam_model, newdata = habitat_grid)
    pred_df <- data.frame(habitat_cover = habitat_grid$habitat_cover, pred = pred_vals)
    
    # Generate object names
    safe_mov <- gsub("[^a-zA-Z0-9]", "", tolower(mov))
    pred_name <- paste0(dataset_name, "_", safe_mov, "_", abil, "_pred")
    gam_name  <- paste0(dataset_name, "_", safe_mov, "_", abil, "_gam")
    data_name <- paste0(dataset_name, "_", safe_mov, "_", abil)
    
    # Store
    results[[pred_name]] <- pred_df
    results[[gam_name]]  <- gam_model
    results[[data_name]] <- df_subset
  }
  
  return(results)
}

delta_results <- process_delta_dataset(df = delta_df, dataset_name = "delta", habitat_grid = habitat_seq)


# Calculate GAM differences -----------------------------------------------

compare_all_predicted_gams <- function(all_results, delta_preds) {
  all_comparisons <- list()
  
  # Filter only *_pred entries
  preds_main <- all_results[grepl("_pred$", names(all_results))]
  preds_delta <- delta_preds[grepl("_pred$", names(delta_preds))]
  
  for (name_main in names(preds_main)) {
    df1 <- preds_main[[name_main]]
    source1_clean <- gsub("_pred$", "", name_main)
    
    for (name_delta in names(preds_delta)) {
      df2 <- preds_delta[[name_delta]]
      source2_clean <- gsub("_pred$", "", name_delta)
      
      # Ensure matching habitat_cover grid
      if (!all.equal(df1$habitat_cover, df2$habitat_cover)) {
        warning(paste("Skipping mismatched habitat_cover between", name_main, "and", name_delta))
        next
      }
      
      # Compute difference and ratio on the original scale
      comparison_df <- data.frame(
        habitat_cover = df1$habitat_cover,
        source_1 = source1_clean,
        source_2 = source2_clean,
        abs_diff = df1$pred - df2$pred,
        abs_ratio = df1$pred / df2$pred,
        stringsAsFactors = FALSE
      )
      
      all_comparisons[[paste0(source1_clean, "_vs_", source2_clean)]] <- comparison_df
    }
  }
  
  final_df <- do.call(rbind, all_comparisons)
  rownames(final_df) <- NULL
  return(final_df)
}


diff_results <- compare_all_predicted_gams(all_results, delta_results)


# Plot GAM differences ----------------------------------------------------

plot_diff_vs_habitat <- function(diff_df, source_group, type, mag, delta_prefix = "delta_") {
  df_sub <- diff_df[diff_df$source_1 %in% source_group & grepl(delta_prefix, diff_df$source_2), ]
  
  # Human-readable labels
  type_label <- switch(type,
                       "ha" = "Low Aggregation",
                       "ma" = "Medium Aggregation",
                       "la" = "High Aggregation",
                       type)
  mag_label <- paste0(mag, "% Movement Increase")
  
  ggplot(df_sub, aes(x = habitat_cover, y = abs_diff, color = source_2)) +
    geom_line(alpha = 0.8) +
    labs(
      title = paste("Absolute Difference in Metapopulation Capacity –", type_label, "–", mag_label),
      x = "Habitat Cover",
      y = "Absolute Difference",
      color = "Delta Scenario"
    ) +
    theme_minimal() +
    facet_wrap(~ source_1, ncol = 3, scales = "free_y") +
    theme(legend.position = "bottom")
}


# Extract unique source_1 names
unique_sources <- unique(diff_results$source_1)

# Get combinations of aggregation type and movement ability
types <- c("ha", "ma", "la")
magnitudes <- c("50", "500", "5000")

plot_list <- list()

for (type in types) {
  for (mag in magnitudes) {
    pattern <- paste0("^", type, "_.*_", mag, "_")
    group_sources <- unique_sources[grepl(pattern, unique_sources)]
    
    if (length(group_sources) == 0) next
    
    p <- plot_diff_vs_habitat(diff_results, source_group = group_sources, type = type, mag = mag)
    print(p)
    plot_list[[paste0(type, "_", mag)]] <- p
  }
}



```


