---
title: "Results"
output: 
  html_document: 
    toc: true
    toc_float: true
    keep_md: true
    number_sections: true
date: "2025-04-24"
---
Return to Model README: [README](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/README.md)

# Exploring Metapopulation Capacity
## The independent effects of habitat cover on metapopulation capacity

Experimental Design: [The independent effects of habitat cover on metapopulation capacity - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-independent-effects-of-habitat-cover-on-metapopulation-capacity)

Changing the size of habitat patches, without changing the distance between the habitat patches gives the following relationship between metapopulation capacity and habitat cover:
```{r area_landscapes_results, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300}


rm(list = ls())

library(tidyverse)
library(patchwork)


# Generate and Plot Habitat Patches ---------------------------------------


# Define 10 equally spaced habitat patches
points_10 <- data.frame(
  x = seq(0, 100, length.out = 10),
  y = seq(0, 100, length.out = 10)
)

# Define 5 equally spaced habitat patches
points_5 <- data.frame(
  x = seq(0, 100, length.out = 5),
  y = seq(0, 100, length.out = 5)
)

# Define 2 equally spaced habitat patches
points_2 <- data.frame(
  x = seq(0, 100, length.out = 2),
  y = seq(0, 100, length.out = 2)
)



# Calculate Metapopulation Capacity ---------------------------------------

# Set alpha
alpha <- 0.05  # change as needed

distance_matrix_2 <- as.matrix(dist(points_2))
distance_matrix_5 <- as.matrix(dist(points_5))
distance_matrix_10 <- as.matrix(dist(points_10))

# Number of patches
n_patches_2 <- nrow(points_2)
n_patches_5 <- nrow(points_5)
n_patches_10 <- nrow(points_10)

# Create results dataframe
results_2 <- data.frame(
  habitat_area = numeric(),
  metapop_cap = numeric()
)
results_5 <- data.frame(
  habitat_area = numeric(),
  metapop_cap = numeric()
)
results_10 <- data.frame(
  habitat_area = numeric(),
  metapop_cap = numeric()
)


# Loop from habitat area 1 to 500 for 2 patches
for (a in 1:500) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_2, ncol = n_patches_2)
  
  dispersal_matrix <- exp(-alpha * distance_matrix_2)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # update results
  results_2 <- rbind(results_2, data.frame(habitat_area = a, metapop_cap = leading_eig))
}


# Loop from habitat area 1 to 500 for 5 patches
for (a in 1:500) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_5, ncol = n_patches_5)
  
  dispersal_matrix <- exp(-alpha * distance_matrix_5)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # update results
  results_5 <- rbind(results_5, data.frame(habitat_area = a, metapop_cap = leading_eig))
}



# Loop from habitat area 1 to 500 for 10 patches
for (a in 1:500) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_10, ncol = n_patches_10)
  
  
  dispersal_matrix <- exp(-alpha * distance_matrix_10)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix# Step 3: Connectivity matrix = area * dispersal
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # update results
  results_10 <- rbind(results_10, data.frame(habitat_area = a, metapop_cap = leading_eig))
}







# Plot Graphs -------------------------------------------------------------

results_2$habitat_cover <- ((results_2$habitat_area * 2)/(100*100)) *100
results_5$habitat_cover <- ((results_5$habitat_area * 5)/(100*100)) *100
results_10$habitat_cover <- ((results_10$habitat_area * 10)/(100*100)) *100


# Plot Metapopulation Capacity

mp_1 <- ggplot(results_2, aes(x = habitat_cover, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_2 <- ggplot(results_5, aes(x = habitat_cover, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_3 <- ggplot(results_10, aes(x = habitat_cover, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )
  

mp_1 / mp_2 / mp_3


```

The log of metapopulation capacity plotted against habitat cover gives the following results:

```{r area_landscapes_log_results, echo=FALSE, fig.height=8,  message=FALSE, warning=FALSE, dpi=300}


# Plot log metapopulation capacity

log_mp_1 <- ggplot(results_2, aes(x = habitat_cover, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_2 <- ggplot(results_5, aes(x = habitat_cover, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_3 <- ggplot(results_10, aes(x = habitat_cover, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Habitat Cover (%)",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )


log_mp_1 / log_mp_2 / log_mp_3



```

## The independent effects of inter-patch distance on metapopulation capacity 

Experimental Design: [The independent effects of habitat cover on metapopulation capacity - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-independent-effects-of-inter-patch-distance-on-metapopulation-capacity)

Scaling the dispersal matrix whilst keeping the habitat area constant gives the following relationship between metapopulation capacity and mean distance between habitat patches.
```{r dispersal_landscapes_results, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300}



rm(list = ls())

library(tidyverse)
library(patchwork)


# Generate and Plot Habitat Patches ---------------------------------------


# Define 10 equally spaced habitat patches
points_10 <- data.frame(
  x = seq(0, 100, length.out = 10),
  y = seq(0, 100, length.out = 10)
)

# Define 5 equally spaced habitat patches
points_5 <- data.frame(
  x = seq(0, 100, length.out = 5),
  y = seq(0, 100, length.out = 5)
)

# Define 2 equally spaced habitat patches
points_2 <- data.frame(
  x = seq(0, 100, length.out = 2),
  y = seq(0, 100, length.out = 2)
)



# Calculate Metapopulation Capacity ---------------------------------------

# Set alpha
alpha <- 0.05  

# Number of patches
n_patches_2 <- nrow(points_2)
n_patches_5 <- nrow(points_5)
n_patches_10 <- nrow(points_10)

distance_matrix_2 <- as.matrix(dist(points_2))
distance_matrix_5 <- as.matrix(dist(points_5))
distance_matrix_10 <- as.matrix(dist(points_10))

# Create results dataframe
results_2 <- data.frame(
  mean_distance = numeric(),
  metapop_cap = numeric()
)
results_5 <- data.frame(
  mean_distance = numeric(),
  metapop_cap = numeric()
)
results_10 <- data.frame(
  mean_distance = numeric(),
  metapop_cap = numeric()
)

# set area to 100 units
a <- 100

# Loop from distance multiplier of 1 to 500
for (d in seq(1, 5, by = 0.01)) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_2, ncol = n_patches_2)
  
  scaled_distance_matrix <- distance_matrix_2 *d
  
  dispersal_matrix <- exp(-alpha * scaled_distance_matrix)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # calculate mean distance
  dist_vector <- scaled_distance_matrix[lower.tri(scaled_distance_matrix)]
  mean_distance <- mean(dist_vector)
  
  # update results
  results_2 <- rbind(results_2, data.frame(mean_distance = mean_distance, metapop_cap = leading_eig))
}


# Loop from distance multiplier of 1 to 500
for (d in seq(1, 5, by = 0.01)) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_5, ncol = n_patches_5)
  
  distance_matrix <- distance_matrix_5 *d
  
  dispersal_matrix <- exp(-alpha * distance_matrix)
  
  diag(dispersal_matrix) <- 0 # set diagonal back to 0
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # calculate mean distance
  dist_vector <- distance_matrix[lower.tri(distance_matrix)]
  mean_distance <- mean(dist_vector)
  
  # update results
  results_5 <- rbind(results_5, data.frame(mean_distance = mean_distance, metapop_cap = leading_eig))
}




# Loop from distance multiplier of 1 to 500
for (d in seq(1, 5, by = 0.01)) {
  
  # generate area matrix
  area_matrix <- matrix(a^2, nrow = n_patches_10, ncol = n_patches_10)
  
  distance_matrix <- distance_matrix_10 *d
  
  dispersal_matrix <- exp(-alpha * distance_matrix)
  
  # multiply dispersal matrix by area matrix
  connectivity_matrix <- area_matrix * dispersal_matrix
  
  # extract eigenvalues
  eig <- eigen(connectivity_matrix)
  
  # find the leading eigenvalue
  leading_eig <- eig$values[1]
  
  # calculate mean distance
  dist_vector <- distance_matrix[lower.tri(distance_matrix)]
  mean_distance <- mean(dist_vector)
  
  # update results
  results_10 <- rbind(results_10, data.frame(mean_distance = mean_distance, metapop_cap = leading_eig))
}







# Plot Graphs -------------------------------------------------------------


# Plot Metapopulation Capacity

mp_1 <- ggplot(results_2, aes(x = mean_distance, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_2 <- ggplot(results_5, aes(x = mean_distance, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

mp_3 <- ggplot(results_10, aes(x = mean_distance, y = metapop_cap)) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )


mp_1 / mp_2 / mp_3

```


The log of metapopulation capacity plotted against mean distance between patches gives the following results:

```{r distance_landscapes_log_results, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE, dpi=300}

# Plot log metapopulation capacity

log_mp_1 <- ggplot(results_2, aes(x = mean_distance, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "2 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_2 <- ggplot(results_5, aes(x = mean_distance, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "5 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )

log_mp_3 <- ggplot(results_10, aes(x = mean_distance, y = log(metapop_cap))) +
  geom_line() +
  labs(
    title = "10 Habitat Patches",
    x = "Mean Distance between Patches",
    y = "Log Metapopulation Capacity"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 10, face = "plain")
  )


log_mp_1 / log_mp_2 / log_mp_3



```

## The effects of habitat configuration on patch distances and areas

Experimental Design: [The effects of habitat configuration on patch distances and areas - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-effects-of-habitat-configuration-on-patch-distances-and-areas)

```{r config_distance_results, echo=FALSE, message=FALSE, warning=FALSE, dpi=300}

rm(list = ls())

library(tidyverse)
library(cowplot)
library(patchwork)

df <- read.csv("Results_Folder/Exploring Metapopulation Capacity/Configuration Distance Relationship/config_dist_summary.csv")

p1 <- ggplot(df, aes(x = as.factor(p), y = edge_density)) +
  geom_boxplot() +
  labs(x = "p", y = "Edge Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p2 <- ggplot(df, aes(x = as.factor(p), y = mean_distance )) +
  geom_boxplot() +
  labs(x = "p", y = "Mean inter-patch Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p3 <- ggplot(df, aes(x = as.factor(p), y = mean_nearest_distance )) +
  geom_boxplot() +
  labs(x = "p", y = "Mean nearest-neighbour Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p4 <- ggplot(df, aes(x = as.factor(p), y = num_patches )) +
  geom_boxplot() +
  labs(x = "p", y = "Number of Patches") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p5 <- ggplot(df, aes(x = as.factor(p), y = mean_patch_area )) +
  geom_boxplot() +
  labs(x = "p", y = "Mean Habitat Patch Area") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p1 + p2 + p3 + p4 + p5


```


## The effects of habitat cover on patch distances and areas

Experimental Design: [The effects of habitat cover on patch distances and areas - Methods](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/Experimental-Design.md#the-effects-of-habitat-cover-on-patch-distances-and-areas)

```{r cover_distance_results, echo=FALSE, message=FALSE, warning=FALSE, dpi=300}


rm(list = ls())

library(tidyverse)
library(cowplot)
library(patchwork)

df <- read.csv("Results_Folder/Exploring Metapopulation Capacity/Cover Distance Relationship/cover_dist_summary.csv")

p1 <- ggplot(df, aes(x = as.factor(ai), y = edge_density)) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Edge Density") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p2 <- ggplot(df, aes(x = as.factor(ai), y = mean_distance )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Mean inter-patch Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p3 <- ggplot(df, aes(x = as.factor(ai), y = mean_nearest_distance )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Mean nearest-neighbour Distance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p4 <- ggplot(df, aes(x = as.factor(ai), y = num_patches )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Number of Patches") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p5 <- ggplot(df, aes(x = as.factor(ai), y = mean_patch_area )) +
  geom_boxplot() +
  labs(x = "Habitat Cover", y = "Mean Habitat Patch Area") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_cowplot()

p1 + p2 + p3 + p4 + p5


```


# The effect of Matrix Quality on Metapopulation Capacity and Persistence

## The effect of Matrix Quality on Metapopulation Capacity

```{r matrix_quality_data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rm(list=ls())

library(tidyverse)

# Import Data -------------------------------------------------------------

# High aggregation group
ha_vhd <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_vhd.csv")
ha_hd  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_hd.csv")
ha_md  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_md.csv")
ha_ld  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_ld.csv")
ha_vld <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ha_csv/ha_vld.csv")

# Medium aggregation group
ma_vhd <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_vhd.csv")
ma_hd  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_hd.csv")
ma_md  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_md.csv")
ma_ld  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_ld.csv")
ma_vld <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/ma_csv/ma_vld.csv")

# Low aggregation group
la_vhd <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_vhd.csv")
la_hd  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_hd.csv")
la_md  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_md.csv")
la_ld  <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_ld.csv")
la_vld <- read.csv("Results_Folder/Matrix Permeability and Metapopulation Capacity/Model 1/la_csv/la_vld.csv")






# Plot Metapopulation Capacities ------------------------------------------

library(ggplot2)
library(grid)
library(patchwork)

# Function for plotting graphs

plot_metapop_vs_cover <- function(df, df_name) {
  p <- ggplot(df, aes(x = habitat_cover, y = log(metapop_cap), colour = movement)) +
    geom_point() +
    geom_smooth(method = "gam", se = FALSE) +
    labs(
      x = "Habitat Cover",
      y = "log Metapopulation Capacity",
      color = "Movement Factor"
    ) +
    theme_minimal()
  
  # Create name like ha_vhd_plot and assign plot object to it in global environment
  plot_name <- paste0(df_name, "_plot")
  assign(plot_name, p, envir = .GlobalEnv)
}


# Function for dispersal labels

dispersal_label <- function(df_name) {
  level_code <- strsplit(df_name, "_")[[1]][2]
  switch(
    level_code,
    "vhd" = "Very High Dispersal",
    "hd" = "High Dispersal",
    "md" = "Medium Dispersal",
    "ld" = "Low Dispersal",
    "vld" = "Very Low Dispersal",
    level_code  # fallback
  )
}

```


### Landscapes with high levels of habitat aggregation

High levels of habitat aggregation with a maximum dispersal distance increase of 50%
```{r matrix_quality_ha_50, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}
# Plot for 50% increase ---------------------------------------------------

# Filter for 50% increases
ha_vhd_50 <- ha_vhd[grepl("50$", as.character(ha_hd$movement)), ]
ha_hd_50 <- ha_hd[grepl("50$", as.character(ha_hd$movement)), ]
ha_md_50 <- ha_md[grepl("50$", as.character(ha_hd$movement)), ]
ha_ld_50 <- ha_ld[grepl("50$", as.character(ha_hd$movement)), ]
ha_vld_50 <- ha_vld[grepl("50$", as.character(ha_hd$movement)), ]


plot_metapop_vs_cover(ha_vhd_50, "ha_vhd_50")
plot_metapop_vs_cover(ha_hd_50, "ha_hd_50")
plot_metapop_vs_cover(ha_md_50, "ha_md_50")
plot_metapop_vs_cover(ha_ld_50, "ha_ld_50")
plot_metapop_vs_cover(ha_vld_50, "ha_vld_50")



# Create labeled plots
ha_vhd_labeled_50 <- ha_vhd_50_plot + ggtitle(dispersal_label("ha_vhd"))
ha_hd_labeled_50 <- ha_hd_50_plot + ggtitle(dispersal_label("ha_hd"))
ha_md_labeled_50 <- ha_md_50_plot + ggtitle(dispersal_label("ha_md"))
ha_ld_labeled_50 <- ha_ld_50_plot + ggtitle(dispersal_label("ha_ld"))
ha_vld_labeled_50 <- ha_vld_50_plot + ggtitle(dispersal_label("ha_vld"))


# Combine them with shared legend
combined_plot <- (
  ha_vhd_labeled_50 + ha_hd_labeled_50 + ha_md_labeled_50
) / (
  ha_ld_labeled_50 + ha_vld_labeled_50 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot

```

High levels of habitat aggregation with a maximum dispersal distance increase of 500%
```{r matrix_quality_ha_500, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 500% increase ---------------------------------------------------

# Filter for 500% increases
ha_vhd_500 <- ha_vhd[grepl("500$", as.character(ha_hd$movement)), ]
ha_hd_500 <- ha_hd[grepl("500$", as.character(ha_hd$movement)), ]
ha_md_500 <- ha_md[grepl("500$", as.character(ha_hd$movement)), ]
ha_ld_500 <- ha_ld[grepl("500$", as.character(ha_hd$movement)), ]
ha_vld_500 <- ha_vld[grepl("500$", as.character(ha_hd$movement)), ]


plot_metapop_vs_cover(ha_vhd_500, "ha_vhd_500")
plot_metapop_vs_cover(ha_hd_500, "ha_hd_500")
plot_metapop_vs_cover(ha_md_500, "ha_md_500")
plot_metapop_vs_cover(ha_ld_500, "ha_ld_500")
plot_metapop_vs_cover(ha_vld_500, "ha_vld_500")



# Create labeled plots
ha_vhd_labeled_500 <- ha_vhd_500_plot + ggtitle(dispersal_label("ha_vhd"))
ha_hd_labeled_500 <- ha_hd_500_plot + ggtitle(dispersal_label("ha_hd"))
ha_md_labeled_500 <- ha_md_500_plot + ggtitle(dispersal_label("ha_md"))
ha_ld_labeled_500 <- ha_ld_500_plot + ggtitle(dispersal_label("ha_ld"))
ha_vld_labeled_500 <- ha_vld_500_plot + ggtitle(dispersal_label("ha_vld"))


# Combine them with shared legend
combined_plot <- (
  ha_vhd_labeled_500 + ha_hd_labeled_500 + ha_md_labeled_500
) / (
  ha_ld_labeled_500 + ha_vld_labeled_500 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot

```

High levels of habitat aggregation with a maximum dispersal distance increase of 5000%
```{r matrix_quality_ha_5000, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 5000% increase ---------------------------------------------------

# Filter for 5000% increases
ha_vhd_5000 <- ha_vhd[grepl("5000$", as.character(ha_hd$movement)), ]
ha_hd_5000 <- ha_hd[grepl("5000$", as.character(ha_hd$movement)), ]
ha_md_5000 <- ha_md[grepl("5000$", as.character(ha_hd$movement)), ]
ha_ld_5000 <- ha_ld[grepl("5000$", as.character(ha_hd$movement)), ]
ha_vld_5000 <- ha_vld[grepl("5000$", as.character(ha_hd$movement)), ]


plot_metapop_vs_cover(ha_vhd_5000, "ha_vhd_5000")
plot_metapop_vs_cover(ha_hd_5000, "ha_hd_5000")
plot_metapop_vs_cover(ha_md_5000, "ha_md_5000")
plot_metapop_vs_cover(ha_ld_5000, "ha_ld_5000")
plot_metapop_vs_cover(ha_vld_5000, "ha_vld_5000")



# Create labeled plots
ha_vhd_labeled_5000 <- ha_vhd_5000_plot + ggtitle(dispersal_label("ha_vhd"))
ha_hd_labeled_5000 <- ha_hd_5000_plot + ggtitle(dispersal_label("ha_hd"))
ha_md_labeled_5000 <- ha_md_5000_plot + ggtitle(dispersal_label("ha_md"))
ha_ld_labeled_5000 <- ha_ld_5000_plot + ggtitle(dispersal_label("ha_ld"))
ha_vld_labeled_5000 <- ha_vld_5000_plot + ggtitle(dispersal_label("ha_vld"))


# Combine them with shared legend
combined_plot <- (
  ha_vhd_labeled_5000 + ha_hd_labeled_5000 + ha_md_labeled_5000
) / (
  ha_ld_labeled_5000 + ha_vld_labeled_5000 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot


```

### Landscapes with medium levels of habitat aggregation

Medium levels of habitat aggregation with a maximum dispersal distance increase of 50%
```{r matrix_quality_ma_50, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 50% increase ---------------------------------------------------

# Filter for 50% increases
ma_vhd_50 <- ma_vhd[grepl("50$", as.character(ma_hd$movement)), ]
ma_hd_50 <- ma_hd[grepl("50$", as.character(ma_hd$movement)), ]
ma_md_50 <- ma_md[grepl("50$", as.character(ma_hd$movement)), ]
ma_ld_50 <- ma_ld[grepl("50$", as.character(ma_hd$movement)), ]
ma_vld_50 <- ma_vld[grepl("50$", as.character(ma_hd$movement)), ]


plot_metapop_vs_cover(ma_vhd_50, "ma_vhd_50")
plot_metapop_vs_cover(ma_hd_50, "ma_hd_50")
plot_metapop_vs_cover(ma_md_50, "ma_md_50")
plot_metapop_vs_cover(ma_ld_50, "ma_ld_50")
plot_metapop_vs_cover(ma_vld_50, "ma_vld_50")



# Create labeled plots
ma_vhd_labeled_50 <- ma_vhd_50_plot + ggtitle(dispersal_label("ma_vhd"))
ma_hd_labeled_50 <- ma_hd_50_plot + ggtitle(dispersal_label("ma_hd"))
ma_md_labeled_50 <- ma_md_50_plot + ggtitle(dispersal_label("ma_md"))
ma_ld_labeled_50 <- ma_ld_50_plot + ggtitle(dispersal_label("ma_ld"))
ma_vld_labeled_50 <- ma_vld_50_plot + ggtitle(dispersal_label("ma_vld"))


# Combine them with smared legend
combined_plot <- (
  ma_vhd_labeled_50 + ma_hd_labeled_50 + ma_md_labeled_50
) / (
  ma_ld_labeled_50 + ma_vld_labeled_50 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot


```

Medium levels of habitat aggregation with a maximum dispersal distance increase of 500%
```{r matrix_quality_ma_500, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}


# Plot for 500% increase ---------------------------------------------------

# Filter for 500% increases
ma_vhd_500 <- ma_vhd[grepl("500$", as.character(ma_hd$movement)), ]
ma_hd_500 <- ma_hd[grepl("500$", as.character(ma_hd$movement)), ]
ma_md_500 <- ma_md[grepl("500$", as.character(ma_hd$movement)), ]
ma_ld_500 <- ma_ld[grepl("500$", as.character(ma_hd$movement)), ]
ma_vld_500 <- ma_vld[grepl("500$", as.character(ma_hd$movement)), ]


plot_metapop_vs_cover(ma_vhd_500, "ma_vhd_500")
plot_metapop_vs_cover(ma_hd_500, "ma_hd_500")
plot_metapop_vs_cover(ma_md_500, "ma_md_500")
plot_metapop_vs_cover(ma_ld_500, "ma_ld_500")
plot_metapop_vs_cover(ma_vld_500, "ma_vld_500")



# Create labeled plots
ma_vhd_labeled_500 <- ma_vhd_500_plot + ggtitle(dispersal_label("ma_vhd"))
ma_hd_labeled_500 <- ma_hd_500_plot + ggtitle(dispersal_label("ma_hd"))
ma_md_labeled_500 <- ma_md_500_plot + ggtitle(dispersal_label("ma_md"))
ma_ld_labeled_500 <- ma_ld_500_plot + ggtitle(dispersal_label("ma_ld"))
ma_vld_labeled_500 <- ma_vld_500_plot + ggtitle(dispersal_label("ma_vld"))


# Combine them with smared legend
combined_plot <- (
  ma_vhd_labeled_500 + ma_hd_labeled_500 + ma_md_labeled_500
) / (
  ma_ld_labeled_500 + ma_vld_labeled_500 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot



```

Medium levels of habitat aggregation with a maximum dispersal distance increase of 5000%
```{r matrix_quality_ma_5000, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}


# Plot for 5000% increase ---------------------------------------------------

# Filter for 5000% increases
ma_vhd_5000 <- ma_vhd[grepl("5000$", as.character(ma_hd$movement)), ]
ma_hd_5000 <- ma_hd[grepl("5000$", as.character(ma_hd$movement)), ]
ma_md_5000 <- ma_md[grepl("5000$", as.character(ma_hd$movement)), ]
ma_ld_5000 <- ma_ld[grepl("5000$", as.character(ma_hd$movement)), ]
ma_vld_5000 <- ma_vld[grepl("5000$", as.character(ma_hd$movement)), ]


plot_metapop_vs_cover(ma_vhd_5000, "ma_vhd_5000")
plot_metapop_vs_cover(ma_hd_5000, "ma_hd_5000")
plot_metapop_vs_cover(ma_md_5000, "ma_md_5000")
plot_metapop_vs_cover(ma_ld_5000, "ma_ld_5000")
plot_metapop_vs_cover(ma_vld_5000, "ma_vld_5000")



# Create labeled plots
ma_vhd_labeled_5000 <- ma_vhd_5000_plot + ggtitle(dispersal_label("ma_vhd"))
ma_hd_labeled_5000 <- ma_hd_5000_plot + ggtitle(dispersal_label("ma_hd"))
ma_md_labeled_5000 <- ma_md_5000_plot + ggtitle(dispersal_label("ma_md"))
ma_ld_labeled_5000 <- ma_ld_5000_plot + ggtitle(dispersal_label("ma_ld"))
ma_vld_labeled_5000 <- ma_vld_5000_plot + ggtitle(dispersal_label("ma_vld"))


# Combine them with smared legend
combined_plot <- (
  ma_vhd_labeled_5000 + ma_hd_labeled_5000 + ma_md_labeled_5000
) / (
  ma_ld_labeled_5000 + ma_vld_labeled_5000 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot


```



### Landscapes with medium levels of habitat aggregation

Low levels of habitat aggregation with a maximum dispersal distance increase of 50%
```{r matrix_quality_la_50, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Low Aggregation Plots --------------------------------------------------


# Plot for 50% increase ---------------------------------------------------

# Filter for 50% increases
la_vhd_50 <- la_vhd[grepl("50$", as.character(la_hd$movement)), ]
la_hd_50 <- la_hd[grepl("50$", as.character(la_hd$movement)), ]
la_md_50 <- la_md[grepl("50$", as.character(la_hd$movement)), ]
la_ld_50 <- la_ld[grepl("50$", as.character(la_hd$movement)), ]
la_vld_50 <- la_vld[grepl("50$", as.character(la_hd$movement)), ]


plot_metapop_vs_cover(la_vhd_50, "la_vhd_50")
plot_metapop_vs_cover(la_hd_50, "la_hd_50")
plot_metapop_vs_cover(la_md_50, "la_md_50")
plot_metapop_vs_cover(la_ld_50, "la_ld_50")
plot_metapop_vs_cover(la_vld_50, "la_vld_50")



# Create labeled plots
la_vhd_labeled_50 <- la_vhd_50_plot + ggtitle(dispersal_label("la_vhd"))
la_hd_labeled_50 <- la_hd_50_plot + ggtitle(dispersal_label("la_hd"))
la_md_labeled_50 <- la_md_50_plot + ggtitle(dispersal_label("la_md"))
la_ld_labeled_50 <- la_ld_50_plot + ggtitle(dispersal_label("la_ld"))
la_vld_labeled_50 <- la_vld_50_plot + ggtitle(dispersal_label("la_vld"))


# Combine them with slared legend
combined_plot <- (
  la_vhd_labeled_50 + la_hd_labeled_50 + la_md_labeled_50
) / (
  la_ld_labeled_50 + la_vld_labeled_50 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot



```

Low levels of habitat aggregation with a maximum dispersal distance increase of 500%
```{r matrix_quality_la_500, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}

# Plot for 500% increase ---------------------------------------------------

# Filter for 500% increases
la_vhd_500 <- la_vhd[grepl("500$", as.character(la_hd$movement)), ]
la_hd_500 <- la_hd[grepl("500$", as.character(la_hd$movement)), ]
la_md_500 <- la_md[grepl("500$", as.character(la_hd$movement)), ]
la_ld_500 <- la_ld[grepl("500$", as.character(la_hd$movement)), ]
la_vld_500 <- la_vld[grepl("500$", as.character(la_hd$movement)), ]


plot_metapop_vs_cover(la_vhd_500, "la_vhd_500")
plot_metapop_vs_cover(la_hd_500, "la_hd_500")
plot_metapop_vs_cover(la_md_500, "la_md_500")
plot_metapop_vs_cover(la_ld_500, "la_ld_500")
plot_metapop_vs_cover(la_vld_500, "la_vld_500")



# Create labeled plots
la_vhd_labeled_500 <- la_vhd_500_plot + ggtitle(dispersal_label("la_vhd"))
la_hd_labeled_500 <- la_hd_500_plot + ggtitle(dispersal_label("la_hd"))
la_md_labeled_500 <- la_md_500_plot + ggtitle(dispersal_label("la_md"))
la_ld_labeled_500 <- la_ld_500_plot + ggtitle(dispersal_label("la_ld"))
la_vld_labeled_500 <- la_vld_500_plot + ggtitle(dispersal_label("la_vld"))


# Combine them with slared legend
combined_plot <- (
  la_vhd_labeled_500 + la_hd_labeled_500 + la_md_labeled_500
) / (
  la_ld_labeled_500 + la_vld_labeled_500 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot




```

Low levels of habitat aggregation with a maximum dispersal distance increase of 5000%
```{r matrix_quality_la_5000, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=8, dpi=300}


# Plot for 5000% increase ---------------------------------------------------

# Filter for 5000% increases
la_vhd_5000 <- la_vhd[grepl("5000$", as.character(la_hd$movement)), ]
la_hd_5000 <- la_hd[grepl("5000$", as.character(la_hd$movement)), ]
la_md_5000 <- la_md[grepl("5000$", as.character(la_hd$movement)), ]
la_ld_5000 <- la_ld[grepl("5000$", as.character(la_hd$movement)), ]
la_vld_5000 <- la_vld[grepl("5000$", as.character(la_hd$movement)), ]


plot_metapop_vs_cover(la_vhd_5000, "la_vhd_5000")
plot_metapop_vs_cover(la_hd_5000, "la_hd_5000")
plot_metapop_vs_cover(la_md_5000, "la_md_5000")
plot_metapop_vs_cover(la_ld_5000, "la_ld_5000")
plot_metapop_vs_cover(la_vld_5000, "la_vld_5000")



# Create labeled plots
la_vhd_labeled_5000 <- la_vhd_5000_plot + ggtitle(dispersal_label("la_vhd"))
la_hd_labeled_5000 <- la_hd_5000_plot + ggtitle(dispersal_label("la_hd"))
la_md_labeled_5000 <- la_md_5000_plot + ggtitle(dispersal_label("la_md"))
la_ld_labeled_5000 <- la_ld_5000_plot + ggtitle(dispersal_label("la_ld"))
la_vld_labeled_5000 <- la_vld_5000_plot + ggtitle(dispersal_label("la_vld"))


# Combine them with slared legend
combined_plot <- (
  la_vhd_labeled_5000 + la_hd_labeled_5000 + la_md_labeled_5000
) / (
  la_ld_labeled_5000 + la_vld_labeled_5000 + plot_spacer()
) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# Display
combined_plot

```


