---
title: "Metapopulation Capacity Model User Guide"
output: 
  html_document: 
    toc: true
    toc_float: true
    keep_md: true
    number_sections: true
date: "2025-04-11"
---
Return to Model README: [README](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/README.md)

# Annotated Example R Model
## Prepare Environment
The first section of code clears any previous information stored in the R environment, installs (or updates) and loads the required packages, and starts a system clock which will record the amount of time the model takes to run. The `devtools` package is required to install `NLMR` and `RandomFields` directly from their repositories.

``` {r prepare environment, eval=FALSE}
rm(list = ls()) # clears environment

# General -----------------------------------------------------------------

install.packages("devtools") # install devtools for packages not on CRAN
library(devtools)

# install packages
install.packages(c("tidyverse", "sf", "units", "ggplot2", "gganimate",
                   "igraph", "raster", "landscapemetrics", "units"))

# install packages not on CRAN
devtools::install_github("ropensci/NLMR")
install.packages('RandomFields', repos =
                   'https://predictiveecology.r-universe.dev', type = 'source')

# load packages
library(tidyverse)
library(parallel)
library(NLMR)
library(RandomFields)
library(landscapemetrics)
library(igraph)
library(raster)
library(sf)
library(units)


# Start timer to measure model run speed
start_timer <- Sys.time()
```


## Create and Loop through Dispersal Scenarios
This section first creates a grid of each possible movement scenario. Each movement scenario consists of the shape of the movement response (eg. `linear`, `convex` etc.) and the percentage increase in mean dispersal ability from a relative yield of 0 to a relative yield of 1 (eg. `50`, `1000` etc.). A loop is then started which will loop through each of the unique movement scenarios, the details of which are added to a list of parameters. 
``` {r create dispersal scenarios, eval=FALSE}

# Loop for dispersal scenarios --------------------------------------------

# Parameterises model dispersal scenarios

# Define movement scenarios
# movement can be 'linear', 'convex', or 'concave'
movements <- c('linear', 'concave', 'convex')

# Define maximum movement increases (%)
movement_abilities <- c(50, 100, 500, 1000)

# Create grid of movement scenarios
movement_combos <- expand.grid(movement = movements, movement_ability = movement_abilities)

# Create a new row for null movement
new_row <- data.frame(movement = "none", movement_ability = 0)

# Add the new row to the grid
movement_combos <- rbind(movement_combos, new_row)

# remove unnecessary variables
rm(movement_abilities, movements, new_row)


# set movement as a character
movement_combos$movement <- as.character(movement_combos$movement)

# Create dataframe of slope constants for movement scenarios
m <- c("linear", "linear", "linear", "linear", "concave", "concave", "concave", "concave", "convex", "convex", "convex", "convex")
m_a <- c("50", "100", "500", "1000", "50", "100", "500", "1000", "50", "100", "500", "1000")
s1 <- c(0.5, 1, 5, 10, 0.5, 1, 5, 10, 2.27e-5, 4.54e-5, 2.27e-4, 4.53e-4)
s2 <- c(NA, NA, NA, NA, NA, NA, NA, NA, 0, 0, 0, 0)
a <- c(1, 1, 1, 1, 0.5, 1, 5, 10, 1, 1, 1, 1)

slopes <- data.frame(m, m_a, s1, s2, a)

# remove unnecessary variables
rm(m, m_a, s1, s2, a)


# create final dataset list
result_final_complete <- list()

# creates a for loop for each of the movement combinations
for (j in 1:nrow(movement_combos)) {
  movement <- movement_combos$movement[j]
  movement_ability <- movement_combos$movement_ability[j]
  
  
  # wrap parameters into list for simulation function
  param_movement <- list(
    movement = movement,
    movement_ability = movement_ability
  )
```

## Create Simulation Function
This section defines a function that will perform the simulation. It starts by setting the landscape parameters including the extent, resolution, level of aggregation and productivity target. See [Model Overview - Landscape Creation](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/model_overview.md) for further details.
``` {r start simulation function, eval=FALSE}
 # Simulation Function -----------------------------------------------------
  
  
  # create a function for replicates of movement scenarios
  rep_function <- function(movement, movement_ability) {
    
    
    
    # Set Landscape Parameters ------------------------------------------------
    
    
    resolution <- 1 # set resolution
    
    x_extent <- 30 # set width
    y_extent <- 30 # set height
    
    landscape_config <- 0.4 # level of patch aggregation
    
    ai <- runif(1, min = 0.01, max = 0.45) #  random proportion of landscape that is habitat between species bounds
    
    # Set landscape productivity target
    productivity <- 0.5
```


## Landscape Creation
A landscape is then generated using the random cluster method within the `NLMR` package. Habitat patches, or 'clumps' are identified using the 8-neighbour method, and edge density is calculated. The raster landscape is then coerced to a data frame and patch numbers are given to each unique habitat patch. The landscape generation code is enveloped within a `repeat` loop to ensure there are at least two habitat patches (a requirement for metapopulation capacity). Finally, the exact proportion of habitat cover is calculated.
``` {r landscape creation, eval=FALSE}
# Generate Landscape ------------------------------------------------------
    
    
    ## Habitat Patches ========================================================
    
    
    # create neutral landscape with >1 patch
    
    repeat {
      # generate landscape using NLMR package
      landscape <- nlm_randomcluster(ncol = x_extent, nrow = y_extent, 
                                     resolution = resolution, 
                                     p = landscape_config, 
                                     ai = c((1-ai),ai))
      
      # identify connected patches
      clumps <- clump(landscape, directions = 8)
      
      # calculate edge_density
      edge_density <- as.numeric(lsm_l_ed(landscape, count_boundary = FALSE, directions = 8)[6])
      
      # coerce to data frame
      landscape_df <- rasterToPoints(landscape, spatial = TRUE)
      landscape_df <- as.data.frame(landscape_df)
      colnames(landscape_df) <- c("layer", "x", "y")
      landscape_df$layer <- as.factor(landscape_df$layer)
      
      # add patch numbers to landscape df
      coordinates(landscape_df) <- ~x + y 
      landscape_df$patch <- extract(clumps, landscape_df)
      landscape_df <- as.data.frame(landscape_df)
      
      # Check the maximum patch number
      max_patch_number <- max(landscape_df$patch, na.rm = TRUE)
      
      # If there is more than one patch, exit the loop
      if (max_patch_number > 1) {
        break
      }
    }
    
    
    # calculate exact landscape cover
    landscape_cover <- (sum(landscape_df$layer == 1)/nrow(landscape_df))*100
    
    # remove unnecessary variables
    rm(clumps, landscape, max_patch_number)
```


## Habitat Patch Distances and Areas
The landscape data frame is then converted back to a raster and the crs is set. The raster is then converted to a polygon (vector) and a distance matrix of inter-patch nearest-neighbour distances is created. The 'patch' numbers in the landscape data frame and the distance matrix may not match, and so this is corrected for by renaming the patch numbers to those assigned during the conversion to polygon. The area of each habitat patch is then calculated, and a landscape plot is generated. Finally, mean patch size and mean inter-patch distance is calculated for recording in the final results.
``` {r distances and areas, eval=FALSE}
## Calculate Distances and Areas ==========================================
    
    
    # Calculate distances
    
    # convert dataframe back to raster
    landscape_df_patches <- landscape_df %>%
      dplyr::select(x, y, patch)
    patch_raster <- rasterFromXYZ(landscape_df_patches)
    
    # remove unnecessary variables
    rm(landscape_df_patches) 
    
    # set resolution of raster in m
    res(patch_raster) <- c(1, 1)
    
    # set crs of raster
    crs(patch_raster) <- CRS("+proj=utm +zone=33 +datum=WGS84")
    
    # Convert raster to polygons
    habitat_polygons <- rasterToPolygons(patch_raster, fun = function(x) {x > 0}, dissolve = TRUE)
    habitat_polygons_sf <- st_as_sf(habitat_polygons)
    
    # remove unnecessary variables
    rm(habitat_polygons, patch_raster) 
    
    # Calculate the distance matrix between the edges of polygons
    dist_matrix <- st_distance(habitat_polygons_sf)
    
    # Check if dist_matrix has units and remove if necessary
    if (inherits(dist_matrix, "units")) {
      dist_matrix <- drop_units(dist_matrix)
    } else {
      dist_matrix <- dist_matrix
    }
    
    
    # find new patch numbers based on polygon data
    new_patch_numbers <- habitat_polygons_sf %>%
      st_drop_geometry()
    new_patch_numbers$index <- 1:nrow(new_patch_numbers)
    
    
    # update patch numbers to new numbers
    landscape_df <- landscape_df %>%
      left_join(new_patch_numbers, by = c("patch" = "patch"))
    
    
    # remove unnecessary variables
    rm(new_patch_numbers, habitat_polygons_sf) 
    
    # clean up landscape data frame
    landscape_df <- landscape_df %>%
      dplyr::select(-patch) %>%    # Remove the 'patch' column
      rename(patch = index) 
    
    # patch number as a factor
    landscape_df$patch <- as.factor(landscape_df$patch)
    
    # Calculate habitat patch sizes
    patch_area_df <- landscape_df %>%
      filter(layer == 1) %>%  # Only consider habitat cells
      group_by(patch) %>%
      summarise(area = n())
    
    # view landscape
    ggplot(landscape_df, aes(x = x, y = y, fill = factor(layer))) +
      geom_tile() +
      scale_fill_manual(
        values = c("0" = "orange", "1" = "darkgreen"),
        name = "Layer"
      ) +
      theme_minimal() +
      theme(
        axis.text = element_blank(),   # Hide axis text
        axis.ticks = element_blank(),  # Hide axis ticks
        panel.grid = element_blank()   # Hide grid lines
      )
    
    # calculate mean habitat patch size
    mean_patch_size <- mean(patch_area_df$area)
    
    # calculate average distance between patches
    dist_vector <- dist_matrix[lower.tri(dist_matrix)]
    mean_distance <- mean(dist_vector)
```


## Matrix Effects
The agricultural yield of the non-habitat matrix is calculated as a function of the landscape production target and the habitat cover. This yield is then normalised relative to the landscape production target to give a relative yield. See [Model Overview - Matrix Effects](https://github.com/benjhodgson/metapop_capacity_matrix/blob/main/Documentation/model_overview.md) for explanation. The functions for the three dispersal responses to yield are then created. The correct slope (constants) for the given movement scenario are taken from the data frame of slope constants created in [1.2](# Create and Loop through Dispersal Scenarios). These constants are passed to the correct function for the given movement scenario and the resultant `dispersal_factor` is recorded. Finally, the mean dispersal distances of the species at a relative yield of 0 and 1 are set based on the movement scenario and the dispersal ability of the species, the actual mean dispersal distance of the species for this landscape is calculated, and the `alpha` ($\alpha$) constant is calculated. 

``` {r matrix effects, eval=FALSE}
# Matrix Effects ----------------------------------------------------------
    
    # calculate agricultural yield
    
    yield <- productivity/(1-(landscape_cover/100))
    
    # normalise yield to get relative yield
    
    relative_yield <- (yield-productivity)/(1-productivity)
    
    
    # functions for calculating alpha (dispersal distance)
    
    
    # Linear
    linear_permeability <- function(y, s, a){ # y = yield, s = slope
      disp_factor = (((s*(1-y)) + a)) 
      return(disp_factor)
    }
    
    # Concave 
    conc_permeability <- function(y, s, a){
      disp_factor = (((-(y)^2*s) + 1)+a) 
      return(disp_factor)
    }
    
    # Convex
    conv_permeability <- function(y, s, s2, a){
      disp_factor = (((exp(10*(1-y)))*s)+s2) + a
      return(disp_factor)
    }
    
    
    
    # pick correct constants (slopes) for current movement scenario
    
    slope <- slopes %>% 
      filter(m == movement & m_a == movement_ability)
    slope <- as.numeric(slope[3])
    slope_2 <- slopes %>% 
      filter(m == movement & m_a == movement_ability)
    slope_2 <- as.numeric(slope_2[4])
    a <- slopes %>% 
      filter(m == movement & m_a == movement_ability)
    a <- as.numeric(a[5])
    
    
    # calculate dispersal multiplier
    
    if (movement == "linear") {
      dispersal_factor <- linear_permeability(y = relative_yield, s = slope, a = a)
    } else if (movement == "concave") {
      dispersal_factor <- conc_permeability(y = relative_yield, s = slope, a = a)
    } else if (movement == "convex") {
      dispersal_factor <- conv_permeability(y  = relative_yield, s = slope, s2 = slope_2, a = a)
    } else if (movement == "none") {
      dispersal_factor <- 1  # movement does not change with yield in null scenarios
    } else {
      print("No matching movement")
    }
    
    
    
    # Set model parameters ====================================================
    
    dispersal_extensive <- 8 # set max mean dispersal distance at relative yield = 0
    dispersal_intense <- dispersal_extensive/ (1+(movement_ability/100)) # calculates mean dispersal distance under relative yield = 1
    dispersal <- dispersal_intense * dispersal_factor # set actual mean dispersal distance based on yield
    alpha <- 1/dispersal # sets alpha
```


## Metapopulation Capacity
To calculate metapopulation capacity, the distance matrix is multiplied by negative $\alpha$ and the leading diagonal is set to 0. This matrix is then multiplied by a newly created area matrix. The matrix eigenvalues are extracted and the leading eigenvalue (metapopulation capacity) is isolated. A data frame of model results is then populated and wrapped into a list with the landscape data frame.
``` {r metapopulation capacity, eval=FALSE}
# Metapopulation Capacity Model -------------------------------------------
    
    metapop <- exp(-alpha * dist_matrix) # exponent of -alpha * distances
    diag(metapop) <- 0 # set diagonal back to 0
    
    area_matrix <- outer(patch_area_df$area, patch_area_df$area, FUN = "*")
    # create a matrix of area products
    
    metapop2 <- metapop * area_matrix # multiply metapop by areas of both habitat patches
    
    eig <- eigen(metapop2) # extract eigenvalues
    
    metapop_cap <- eig$values[1] # isolate leading eigenvalue
    
    # populates results
    results <- data.frame(
      metapop_cap = metapop_cap,
      alpha = alpha,
      resolution = resolution,
      x_extent = x_extent,
      y_extent = y_extent,
      landscape_config = landscape_config,
      landscape_cover = landscape_cover,
      edge_density = edge_density,
      mean_distance = mean_distance,
      mean_patch_size = mean_patch_size)
    
    # wraps results and landscape data into a list
    results_list <- list(
      results = results,
      landscape = landscape_df
    )
    
    # returns list of results
    return(results_list)
    
  }
```

## Run Model
The final section of the model loops through each movement scenario and applies the simulation function. The number of replicates (for each movement scenario) is defined and an overall seed is set. Seeds are then generated for each repetition, ensuring landscapes across movement scenarios are the same. The simulation function is then applied in parallel (doesn't work in Windows) using a specified number of cores. The lists of results for each replicate and movement scenario are wrapped into a final nested list of data called `results_final_complete`. The model timer is stopped, the model runtime printed, and the results saved to the local device. 
``` {r run model, eval=FALSE}
# Run Model ---------------------------------------------------------------
  
  num_reps <- 1 # specify the number of repeats
  
  set.seed(123) # sets the overall model seed so that landscapes are consistent between model runs
  
  num_reps <- 2 # specify the number of repeats
  
  seeds <- sample.int(1e6, num_reps) # choose seeds for landscapes
  
  # Set the number of cores for parallel processing
  num_cores <- 1
  
  # Use mclapply to apply the function 100 times in parallel
  result_final <- mclapply(1:num_reps, function(i) {
    
    set.seed(seeds[i]) # set the rep-specific seed
    
    # Apply the function with the parameters passed from the list
    do.call(rep_function, param_movement)
  }, mc.cores = num_cores)
  
  
  result_final_name <- paste("result_final", movement, movement_ability, sep = "_")
  
  assign(result_final_name, result_final)
  
  result_final_complete[[result_final_name]] <- result_final
  
  cat(movement, movement_ability, " ")
  
}


# print duration of program run
print( Sys.time() - start_timer)

# Save outputs ------------------------------------------------------------

saveRDS(result_final_complete, file = "Result_Final")
```


## Data Extraction

## Table of Movement Scenario Constants

